---
output:
  github_document:
    toc: true
    toc_depth: 3
    includes:
      before_body: inst/header.md
always_allow_html: yes
editor_options:
  markdown:
    wrap: sentence
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  fig.path = "man/figures",
  fig.width = 10,
  asciicast_theme = if (Sys.getenv("IN_PKGDOWN") == "true") "pkgdown" else "readme"
)
asciicast::init_knitr_engine(
  echo = TRUE,
  echo_input = FALSE,
  startup = quote({
    library(cli)
    options(cli.num_colors = cli::truecolor)
  })
)
```

```{asciicast asciicast-setup, include = FALSE, results = "hide"}
pkgload::load_all()
# emoji output is slightly incorrect currently, maybe a font issue
options(pkg.emoji = FALSE)
# we do this to have a package to use in the examples
setwd("/tmp")
if (!file.exists("cli")) system("git clone --depth 1 https://github.com/r-lib/cli")
setwd("cli")
unlink(".github/workflows/rhub.yaml")
```

## Installation

Install rhub from CRAN:

```{r, asciicast-install, eval = FALSE, cache = FALSE}
pak::pkg_install("rhub")
```

## Usage

### Requirements

- A Github account.
- Your R package must be in a GitHub repository.
- You need a GitHub [Personal Access Token](https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/creating-a-personal-access-token).
  You can use the [gitcreds package](https://gitcreds.r-lib.org/) to add
  the token to the git credential store.

See the [R Consortium runners](#the-r-consortium-runners) section for
using rhub if your package is not on GitHub.

### Private repositories

rhub uses GitHub Actions, which is free for public repositories.
For private repositories you also get some minutes for free, depending on
the GitHub subscription you have. See
[About billing for GitHub Actions](https://docs.github.com/en/billing/managing-billing-for-github-actions/about-billing-for-github-actions) for details.

### Setup

1. Switch to the directory of your package, and call `rhub::rhub_setup()` to
  add the R-hub workflow file to your package.
```{asciicast rhub-setup}
rhub::rhub_setup()
```

2. Run `git commit` and `git push` to push the workflow file to GitHub.

3. Run `rhub::rhub_doctor()` to check if everything is set up correctly:
```{asciicast rhub-doctor}
rhub::rhub_doctor()
```

### Run checks

Use `rhub::rhub_platforms()` to get a list of supported platforms and checks:
```{asciicast rhub-platforms}
rhub::rhub_platforms()
```

```{asciicast include = FALSE}
testthat::local_mocked_bindings(
  gh_rest_post = function(...) list(status_code = 204L),
  readline = function(prompt) {
    cat(prompt)
    Sys.sleep(1)
    cat("1, 5\n")
    "1, 5"
  }
)
```

Run `rhub::rhub_check()` to start R-hub v2 checks on GitHub Actions:
```{asciicast rhub-check}
rhub::rhub_check()
```

## The R Consortium runners

If you don't want to put your package on GitHub, you can still use the
rhub package to run package checks on any supported platform using a
shared pool of runners in the https://github.com/r-hub2 GitHub
organization.

The process is similar to the first version of R-hub:

* Set your working directory to the R package you want to check.
* Obtain a token from R-hub, to verify your email address:
  ```
  rc_new_token()
  ```
  (You do not need to do this, if you already submitted packages to a
  previous version of R-hub from the same machine, using the same email
  address. Call `rc_list_local_tokens()` to check if you already have
  tokens.)
* Submit a build with
  ```
  rc_submit()
  ```
* Select the platforms you want to use, and follow the instructions and
  the link provided to see your check results.

### Limitations of the R Consortium runners

* You package will be public for the world, and will be stored in the
  https://github.com/r-hub2 organization. Your check output and results
  will be public for anyone with a GitHub account. If you want to keep
  your package private, you can put it in a private GitHub repository,
  and use the `rhub_setup()` and `rhub_check()` functions instead.
* The R Consortium runners are shared among all users, so you might need
  to wait for your builds to start.
* You have to wait at least five minutes between submissions with
  `rc_submit()`.
* Currently you need to create a GitHub account to see the check logs of
  your package. You don't need a GitHub account to submit the checks.

To avoid these limitations (except for the neeed for a GitHub accounr),
put your package in a GitHub repository, and use the `rhub_setup()` and
`rhub_check()` functions instead of `rc_submit()` and the R Consortium
runners.

## Code of Conduct

Please note that the rhub package is released with a
[Contributor Code of Conduct](https://r-hub.github.io/rhub/dev/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.

## License

MIT © R Consortium
