---
output:
  github_document:
    toc: true
    toc_depth: 3
    includes:
      before_body: inst/header.md
always_allow_html: yes
editor_options:
  markdown:
    wrap: sentence
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  fig.path = "man/figures",
  fig.width = 10,
  asciicast_theme = if (Sys.getenv("IN_PKGDOWN") == "true") "pkgdown" else "readme"
)
asciicast::init_knitr_engine(
  echo = TRUE,
  echo_input = FALSE,
  startup = quote({
    library(cli)
    options(cli.num_colors = cli::truecolor)
  })
)
```

```{asciicast asciicast-setup, include = FALSE, results = "hide"}
pkgload::load_all()
# emoji output is slightly incorrect currently, maybe a font issue
options(pkg.emoji = FALSE)
# we do this to have a package to use in the examples
setwd("/tmp")
if (!file.exists("cli")) system("git clone --depth 1 https://github.com/r-lib/cli")
setwd("cli")
unlink(".github/workflows/rhub.yaml")
```

## Installation

Install rhub from CRAN:

```{r, asciicast-install, eval = FALSE, cache = FALSE}
pak::pkg_install("rhub")
```

## Usage

### Requirements

- A Github account.
- Your R package must be in a GitHub repository.
- You need a GitHub [Personal Access Token](https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/creating-a-personal-access-token).
  You can use the [gitcreds package](https://gitcreds.r-lib.org/) to add
  the token to the git credential store.

### Private repositories

rhub uses GitHub Actions, which is free for public repositories.
For private repositories you also get some minutes for free, depending on
the GitHub subscription you have. See
[About billing for GitHub Actions](https://docs.github.com/en/billing/managing-billing-for-github-actions/about-billing-for-github-actions) for details.

### Setup

1. Switch to the directory of your package, and call `rhub::rhub_setup()` to
  add the R-hub workflow file to your package.
```{asciicast rhub-setup}
rhub::rhub_setup()
```

2. Run `git commit` and `git push` to push the workflow file to GitHub.

3. Run `rhub::rhub_doctor()` to check if everything is set up correctly:
```{asciicast rhub-doctor}
rhub::rhub_doctor()
```

### Run checks

Use `rhub::rhub_platforms()` to get a list of supported platforms and checks:
```{asciicast rhub-platforms}
rhub::rhub_platforms()
```

```{asciicast include = FALSE}
testthat::local_mocked_bindings(
  gh_rest_post = function(...) list(status_code = 204L),
  readline = function(prompt) {
    cat(prompt)
    Sys.sleep(1)
    cat("1, 5\n")
    "1, 5"
  }
)
```

Run `rhub::rhub_check()` to start R-hub v2 checks on GitHub Actions:
```{asciicast rhub-check}
rhub::rhub_check()
```

## Code of Conduct

Please note that the callr project is released with a
[Contributor Code of Conduct](https://callr.r-lib.org/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.

## License

MIT © R Consortium
