---
title: "Local Linux checks with Docker"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Local Linux checks with Docker}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rhub)
```

## Introduction

Scenario: there's a bug in the check results of your package on a CRAN Linux
platform, or you saw such a bug even before CRAN submission, by building your
 package on a R-hub Linux platform. How can you reproduce and fix the bug? 
 Submitting to the R-hub platform ([or the R-hub platform that's closest to the CRAN platform](https://docs.r-hub.io/#rhub-cran-platforms)) after each tweak of 
your code would have a high turnaround so is not optimal for debugging. 
Thankfully, R-hub Linux Docker images are available for you to use, so you
can run the R-hub Linux builders locally, and `rhub` features helper functions.

## Install and get to know Docker

To be able to use the feature, you will need to install Docker. Please refer 
to [Docker docs](https://docs.docker.com/install/). On Windows, installation 
might be trickier, check that your machine [meets the system requirements](https://docs.docker.com/docker-for-windows/install/#what-to-know-before-you-install).

If you are new to Docker, this tutorial [feature a nice introduction](https://ropenscilabs.github.io/r-docker-tutorial/01-what-and-why.html).

## List R-hub Linux images

Each of R-hub Linux platforms is associated to a Docker image, whose Dockerfile 
is stored in the [r-hub/rhub-linux-builders repository](https://github.com/r-hub/rhub-linux-builders#rhub-linux-builders), and that is built and 
available on Docker Hub. _Note, if you're used to using Docker images outside of R, you might want to 
just refer to the information in [R-hub Linux Docker images GitHub repository](https://github.com/r-hub/rhub-linux-builders#rhub-linux-builders) (
including links to the built image on Docker Hub)._

To list the available images from R, you can use the `local_check_linux_images()` 
function that returns a `data.frame` and has a pretty default printing.

```{r list}
imgs <- local_check_linux_images()
imgs
knitr::kable(imgs, row.names = FALSE)
```

Of particular interest are

* the `cranname` columns if you're trying to find an equivalent to a CRAN platform;

* the `name` platform which is the ID you should use to select that platform.

## Running local checks

Below we'll start a check of a package on the "debian-gcc-release" image (
Debian Linux, R-release, GCC).

```r
pkg_path <- "/home/maelle/Documents/R-hub/test-packages/note"
rhub::local_check_linux(pkg_path, image = "debian-gcc-release")
```