---
title: "Local Linux checks with Docker"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Local Linux checks with Docker}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rhub)
```

## Introduction

Scenario: there's a bug in the check results of your package on a CRAN
Linux platform, or you saw such a bug even before CRAN submission, by
building your package on a R-hub Linux platform. How can you reproduce and
fix the bug?  Submitting to the R-hub platform
([or the R-hub platform that's closest to the CRAN platform](https://docs.r-hub.io/#rhub-cran-platforms))
after each tweak of your code would have a high turnaround so is not
optimal for debugging. R-hub's Linux Docker images are available for you to
use, so you can run the R-hub Linux builders locally.

**Warning: at the moment, the functions are not tested on Windows! Bug
  reports are welcome :-)** 

## Install and get to know Docker

To be able to use the feature, you will need to install Docker. Please
refer to [Docker docs](https://docs.docker.com/install/). On Windows,
installation might be trickier, check that your machine
[meets the system requirements](https://docs.docker.com/docker-for-windows/install/#what-to-know-before-you-install). On
Linux, make sure to
[run the post-installation steps](https://docs.docker.com/install/linux/linux-postinstall/)
to make the `docker` command available to your user without the `sudo`
prefix. 

If you are new to Docker, for the basic use shown in the next two sections
you don't need to learn anything, you won't have to leave R. Nonetheless,
if you're curious, this tutorial
[features a nice introduction](https://ropenscilabs.github.io/r-docker-tutorial/01-what-and-why.html).
Also see [this blog post](https://colinfay.me/docker-r-reproducibility/)
and the list of resources it shows at the end.

## List R-hub Linux images

Each of R-hub Linux platforms is associated to a Docker image, whose
Dockerfile is stored in the
[r-hub/rhub-linux-builders repository](https://github.com/r-hub/rhub-linux-builders#rhub-linux-builders),
and that is built and available on Docker Hub. Note, if you're used to
using Docker images outside of R, you might want to just refer to the
information in
[R-hub Linux Docker images GitHub repository](https://github.com/r-hub/rhub-linux-builders#rhub-linux-builders)
(including links to the built images on Docker Hub). The advantage of using
the rhub package instead of Docker directly, is that the package will
install the system requirements properly.

To list the available images from R, you can use the
`local_check_linux_images()` function that returns a `data.frame` and has a
pretty default printing. 

```{r list}
imgs <- local_check_linux_images()
imgs
knitr::kable(imgs, row.names = FALSE)
```

Of particular interest are

* the `cranname` columns if you're trying to find an equivalent to a CRAN
  platform; 

* the `name` platform which is the ID you should use to select that
  platform. 

In theory, you could also use images that are not listed in the list above,
e.g. your own Docker images.

## Run local checks

Below we'll start a check of a package on the "rhub/debian-gcc-release"
image (Debian Linux, R-release, GCC). The first time you use an image on
your machine, it'll be downloaded from Docker Hub, which might take a
while. The image won't be deleted after use, so next time will be faster
until you clean up your machine's Docker images, which one should do once
in a while (note that R-hub images are regularly updated).

```r
pkg_path <- "/home/maelle/Documents/R-hub/test-packages/note"
local_check_linux(pkg_path, image = "rhub/debian-gcc-release")
```

You can either just run the check as shown above, which will print a log to
the screen, including `R CMD check` results in the end, or assign it to an
object:

```r
pkg_path <- "/home/maelle/Documents/R-hub/test-packages/note"
chk <- local_check_linux(pkg_path, image = "rhub/debian-gcc-release")
```

The object returned is of the class `rcmdcheck::rcmdcheck` which is an S3
object with fields `errors`, `warnings` and `notes` (character vectors),
that you could operate on if you wish.

The `local_check_linux()` function creates a container (instance of the
image) that won't be deleted after use so you might want to clean up once
in a while.

## Do more with R-hub Linux images

If running checks in images iteratively isn't enough for your debugging,
you might want to run the container created by `local_check_linux()`. Take
note of the container name and run (in a shell, not in R)

```
docker container start 7181196d-bc3c-4fc8-a0e8-dc511150335d-2
docker exec -it 7181196d-bc3c-4fc8-a0e8-dc511150335d-2 bash
```

where `7181196d-bc3c-4fc8-a0e8-dc511150335d-2` is the container name, this
is printed out by `local_check_linux()`. After running these commands, you
will get a shell within the Docker container, where you can run R. Note
that on some containers R is installed in `/opt/`.

For more information, you may want to look at the shell script that `rhub`
uses to set up the container for running the check. To find it, run the
code below.

```{r}
system.file(package = "rhub", "bin", "rhub-linux-docker.sh")
```
